Ground Truth Sampling Cells Colorado Plateau

### 2024-04-03 scripts/GroundTruthSampling.Rmd
summary lyr 3

r <- rast(blm_surf, resolution = 90, nlyrs = 3) # create 30 meter resolution raster
datr <- lapply(dat, rasterize, r, field = 'Rank')
datr <- rast(datr)
r[[1]] <- app(ifel(datr == 1, 1, NA), sum, na.rm = TRUE)
r[[2]] <- app(ifel(datr %in% c(2, 3, 4), 1, NA, na.rm = TRUE))
datr2 <-

rm(dat)
```

>>>>

r <- rast(blm_surf, resolution = 90, nlyrs = 3) # create 30 meter resolution raster
names(r) <- c('SumOccupiedPatches', 'Sum1stOrderNeighboringPatches', 'CommonessInversion')
datr <- lapply(dat, rasterize, r, field = 'Rank')
datr <- rast(datr)

# count the occupied patches by overlap  
r[[1]] <- app(ifel(datr == 1, 1, NA), sum, na.rm = TRUE)

# count the patches which are first order neighbors from a known-populated locality  
r[[2]] <- app(ifel(datr %in% c(2, 3, 4), 1, NA), sum, na.rm = TRUE)

# How common are occupied + 1st order neighbors for each species? i.e. which are less common
# and should be considered more importantly during scouting ??

datr <- ifel(datr <= 4, 1, NA)
tab <- freq(datr)

tab$InverseWt <- 1 - ( tab$count / max(tab$count))
tab$NegativeInverseWt <- round( 1 / (max(tab$InverseWt+1) - tab$InverseWt), 2) * 100
# extreme skew rescale with 'negative transform', keep digits few for memory.

par(mfrow=c(1,2)) 
hist(tab$InverseWt, main = 'Raw', xlab = 'Weight') 
hist(tab$NegativeInverseWt, main = 'Negative Transform', xlab = 'Weight')

for (i in seq_along(1:dim(datr)[3])){ 
  datr[[i]] <- ifel(datr[[i]] == 1, tab$NegativeInverseWt[i], NA) 
}

r[[3]] <- app(datr, sum, na.rm = TRUE)
names(r) <- c('SumOccupiedPatches', 'Sum1stOrderNeighboringPatches', 'SumInverseCommonness')
plot(r)

writeRaster(r, '../results/Summary_Rasters/COPL.tif')

rm(datr, tab, i)
```
